### 2019-03-20 doc/leader/declaration_authorship.tex

\chapter*{Declaration of primary authorship}
\label{chap:declaration_authorship}

\vspace*{10mm}

I / We hereby confirm that I / we have written this thesis independently and without using other sources and resources than those specified in the bibliography. All text passages which were not written by me are marked as quotations and provided with the exact indication of its origin.

\vspace{15mm}

\begin{tabbing}
xxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\kill
Place, Date:		\> [Biel/Burgdorf], \versiondate \\ \\ 
Last Name/s, First Name/s:	\> [Test Peter] 	\> [M\"uster R\"os\"a] \\ \\ \\ \\ 
Signature/s:	\> ......................................\> ...................................... \\
\end{tabbing}