### 2019-03-20 doc/leader/frontpage_with_picture.tex

%
% Project documentation template
% ===========================================================================
% This is part of the document "Project documentation template".
% Authors: brd3, kaa1
%

\begin{titlepage}

% BFH-Logo absolute placed at (28,12) on A4 and picture (16:9 or 15cm x 8.5cm)
% Actually not a realy satisfactory solution but working.
%---------------------------------------------------------------------------
\setlength{\unitlength}{1mm}
\begin{textblock}{20}[0,0](28,12)
	\includegraphics[scale=1.0]{images/BFH_Logo_B.png}
\end{textblock}

\begin{textblock}{154}(28,48)
	\begin{picture}(150,2)
		\put(0,0){\color{bfhgrey}\rule{150mm}{2mm}}
	\end{picture}
\end{textblock}

\begin{textblock}{154}[0,0](28,50)
	\includegraphics[scale=1.0]{images/placemarker.jpg}			% define cover picture
\end{textblock}

\begin{textblock}{154}(28,135)
	\begin{picture}(150,2)
		\put(0,0){\color{bfhgrey}\rule{150mm}{2mm}}
	\end{picture}
\end{textblock}
\color{black}

% Institution / titel / subtitel / authors / experts:
%---------------------------------------------------------------------------
\begin{flushleft}

\vspace*{115mm}

\fontsize{26pt}{28pt}\selectfont 
\heading				\\							% Read heading from file leader/title.tex
\vspace{2mm}

\fontsize{16pt}{20pt}\selectfont\vspace{0.3em}
Place your subheading here 			\\				% Insert subheading
\vspace{5mm}

\fontsize{10pt}{12pt}\selectfont
\textbf{Description of thesis (semester- / Bachelor thesis / etc.)} \\		% Insert text
\vspace{3mm}

% Abstract (eingeben):
%---------------------------------------------------------------------------
\begin{textblock}{150}(28,190)
\fontsize{10pt}{12pt}\selectfont
[Insert short text (abstract) if desired] \\ 
This document serves as a template for the compilation of reports according to the guidelines of the BFH. The template is written in LATEX and supports the automatic writing of various directories, references, indexing and glossaries. This small text is a summary of this document with a length of 4 to max. 8 lines. \\ 
The cover picture may be turned on or off in the lines 157/158 of the file template.tex.
\end{textblock}

\begin{textblock}{150}(28,225)
\fontsize{10pt}{17pt}\selectfont
\begin{tabbing}
xxxxxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx \kill
Degree course:	\> [z.B. Electrical and Communication Engineering]	\\		% insert name of degree course
Authors:		\> [Test Peter, M\"uster R\"os\"a]		\\					% insert names
Tutor:	\> [Dr.~Xxxx Xxxx, Dr.~Yyyy Yyyy]		\\							% insert names
Constituent:	\> [Wwwww AG]					\\							% insert names
Experts:		\> [Dr.~Zzzz Zzzz]				\\							% insert names
Date:			\> \versiondate					\\							% read from file leader/version.tex
\end{tabbing}

\end{textblock}
\end{flushleft}

\begin{textblock}{150}(28,280)
\noindent 
\color{bfhgrey}\fontsize{9pt}{10pt}\selectfont
Berner Fachhochschule | Haute \'ecole sp\'ecialis\'ee bernoise | Bern University of Applied Sciences
\color{black}\selectfont
\end{textblock}

\end{titlepage}

%
% ===========================================================================
% EOF
%