### 2019-05-08 doc/chapters/using_bitcoins.tex
changed top level titles

\chapter{Using Bitcoin-S}
\label{chap:using_bitcoins}

\section{Creating of a transaction (valid, unvalid with TxBuilder)}

\section{Validation of a transaction (with method checkTransaction)}

\section{{Bugfix for noDublicate}}