### 2019-05-14 doc/leader/frontpage_with_picture.tex
titlepage

%
% Project documentation template
% ===========================================================================
% This is part of the document "Project documentation template".
% Authors: brd3, kaa1
%

\begin{titlepage}

% BFH-Logo absolute placed at (28,12) on A4 and picture (16:9 or 15cm x 8.5cm)
% Actually not a realy satisfactory solution but working.
%---------------------------------------------------------------------------
\setlength{\unitlength}{1mm}
\begin{textblock}{20}[0,0](28,12)
	\includegraphics[scale=1.0]{images/BFH_Logo_B.png}
\end{textblock}

\begin{textblock}{154}(28,48)
	\begin{picture}(150,2)
		\put(0,0){\color{bfhgrey}\rule{150mm}{2mm}}
	\end{picture}
\end{textblock}

\begin{textblock}{154}[0,0](28,50)
	\includegraphics[scale=1.0]{images/placemarker.jpg}			% define cover picture
\end{textblock}

\begin{textblock}{154}(28,135)
	\begin{picture}(150,2)
		\put(0,0){\color{bfhgrey}\rule{150mm}{2mm}}
	\end{picture}
\end{textblock}
\color{black}

% Institution / titel / subtitel / authors / experts:
%---------------------------------------------------------------------------
\begin{flushleft}

\vspace*{115mm}

\fontsize{26pt}{28pt}\selectfont 
\heading				\\							% Read heading from file leader/title.tex
\vspace{2mm}

\fontsize{16pt}{20pt}\selectfont\vspace{0.3em}
Place your subheading here 			\\				% Insert subheading
\vspace{5mm}

\fontsize{10pt}{12pt}\selectfont
\textbf{Bachelor thesis} \\		% Insert text
\vspace{3mm}

% Abstract (eingeben):
%---------------------------------------------------------------------------
\begin{textblock}{150}(28,190)
\fontsize{10pt}{12pt}\selectfont
[Insert short text (abstract) if desired] \\ 
This document serves as a template for the compilation of reports according to the guidelines of the BFH. The template is written in LATEX and supports the automatic writing of various directories, references, indexing and glossaries. This small text is a summary of this document with a length of 4 to max. 8 lines. \\ 
The cover picture may be turned on or off in the lines 157/158 of the file template.tex.
\end{textblock}

\begin{textblock}{150}(28,225)
\fontsize{10pt}{17pt}\selectfont
\begin{tabbing}
xxxxxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx \kill
Degree course:	\> Computer Science	\\		% insert name of degree course
Authors:		\> Anna Doukmak, Ramon Boss		\\					% insert names
Tutor:	\> Kai Brünnler @TODO Dr.?		\\							% insert names
%Constituent:	\> [Wwwww AG]					\\							% insert names
Experts:		\> Urs Keller				\\							% insert names
Date:			\> \versiondate					\\							% read from file leader/version.tex
\end{tabbing}

\end{textblock}
\end{flushleft}

\begin{textblock}{150}(28,280)
\noindent 
\color{bfhgrey}\fontsize{9pt}{10pt}\selectfont
Berner Fachhochschule | Haute \'ecole sp\'ecialis\'ee bernoise | Bern University of Applied Sciences
\color{black}\selectfont
\end{textblock}

\end{titlepage}

%
% ===========================================================================
% EOF
%