### 2019-05-22 doc/leader/versions.tex
added some todos and fixed version

% Control of versions :
% -----------------------------------------------

\begin{textblock}{180}(15,150)
\color{black}
\begin{huge}
Versions
\end{huge}
\vspace{10mm}

\fontsize{10pt}{18pt}\selectfont
\begin{tabbing}
xxxxxxxxxxx\=xxxxxxxxxxxxxxx\=xxxxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx \kill
Version	\> Date	\> Status			\> Remarks		\\
1.0	\> TODO fix this	\> Final		\> Final Version 	\\
\end{tabbing}

\end{textblock}