### 2019-05-22 doc/chapters/towards_verifying.tex
changed title

\chapter{Towards verifyng Bitcoin-S with Stainless}
\label{chap:connecting}

\section{Integration (Versionskonflikte, neues Plugin)}

\section{Error Reporting with sbt and JAR}

\section{Trying to Verify checkTransaction}

\subsection{Findings}

\subsection{Bugfix}

\section{Verification of Method "+" // Change this}

\subsection{Rewriting Generics}

\subsection{Rewriting Objects}

\subsection{Rewriting BigInt Constructor (only literal argument, no long argument, etc.)}

\subsection{Rewriting Usage of BigInt "\&"}

\subsection{Rewriting require (funktioniert evtl?)}