### 2019-05-22 doc/chapters/towards_verifying.tex
more text

\chapter{Towards verifyng Bitcoin-S with Stainless}
\label{chap:connecting}

\section{Integration (Versionskonflikte, neues Plugin)}

\section{Error Reporting with sbt and JAR}

\section{Trying to Verify checkTransaction}

\subsection{Findings}

\subsection{Bugfix}

\section{Verification of Method "+ 0" // Change this}
After so many failures, we have decided to search for the smallest unit in Bitcoin-S-Core that is worthwhile to verify.
We found the addition of two Satoshis would be a good candidate.
To make it even more easy, we decided to verify only the addition, where we add zero to an amount of Satoshis.
The signature looks like the following.
\begin{lstlisting}
  +(c: CurrencyUnit): CurrencyUnit
\end{lstlisting}
Where the class Satoshis extends CurrencyUnit.
The post condition would be, that the parameter c must be zero.
\begin{lstlisting}
  require(c.satoshis.underlying == Int64.zero)
\end{lstlisting}
\emph{c.satoshis} is an abstract method of CurrencyUnit that must return an instance of the class Satoshis.

\emph{c.satoshis.underlying} is also an abstract method of CurrencyUnit that must return an instance of the abstract type A.

Both are implemented in Satoshis where A is set to Int64.
So the underlying number of the parameter must be zero.

Next we ensure, that the result is the same value as \emph{this}.
\begin{lstlisting}
  ensuring(res => res.satoshis == this.satoshis)
\end{lstlisting}
Here, we can directly use equals (==) on Satoshis, because its a case class and Int64, the only parameter of Satoshis is a case class too.

\subsection{Rewriting Abstract Type Member}
Stainless output:
\begin{lstlisting}
[ Error  ] CurrencyUnits.scala:5:3: Stainless doesn't support abstract type members
           type A
\end{lstlisting}

\subsection{Rewriting Generics}

\subsection{Rewriting Objects}
Stainless output:
\begin{lstlisting}
[ Error  ] CurrencyUnits.scala:54:1: Objects cannot extend classes or implement
           traits, use a case object instead
           object Satoshis extends BaseNumbers[Satoshis] {
\end{lstlisting}

\subsection{Rewriting BigInt Constructor (only literal argument, no long argument, etc.)}
Stainless output:
\begin{lstlisting}
[ Error  ] CurrencyUnits.scala:47:33: Only literal arguments are allowed for BigInt.
           def toBigInt: BigInt = BigInt(toLong)
\end{lstlisting}

\subsection{Rewriting Usage of BigInt \&-Function}

\subsection{Rewriting require (funktioniert evtl?)}