### 2019-05-23 doc/chapters/towards_verifying.tex
more text

\chapter{Towards verifyng Bitcoin-S with Stainless}
\label{chap:connecting}

\section{Integration (Versionskonflikte, neues Plugin)}

\section{Error Reporting with sbt and JAR}

\section{Trying to Verify checkTransaction}

\subsection{Findings}

\subsection{Bugfix}

\section{Verification of Method "+ 0" // Change this}
After so many failures, we have decided to search for the smallest unit in Bitcoin-S-Core that is worthwhile to verify.
We found the addition of two Satoshis would be a good candidate.
To make it even more easy, we decided to verify only the addition, where we add zero to an amount of Satoshis.
The signature looks like the following.
\begin{lstlisting}
  +(c: CurrencyUnit): CurrencyUnit
\end{lstlisting}
Where the class Satoshis extends CurrencyUnit.
The post condition would be, that the parameter c must be zero.
\begin{lstlisting}
  require(c.satoshis.underlying == Int64.zero)
\end{lstlisting}
\emph{c.satoshis} is an abstract method of CurrencyUnit that must return an instance of the class Satoshis.

\emph{c.satoshis.underlying} is also an abstract method of CurrencyUnit that must return an instance of the abstract type A.

Both are implemented in Satoshis where A is set to Int64.
So the underlying number of the parameter must be zero.

Next we ensure, that the result is the same value as \emph{this}.
\begin{lstlisting}
  ensuring(res => res.satoshis == this.satoshis)
\end{lstlisting}
Here, we can directly use equals (==) on Satoshis, because its a case class and Int64, the only parameter of Satoshis is a case class too.

After writing the verification, we ran Stainless on the source files.
It prints a lot of errors because of code incompatibility.
So we had to rewrite it as described in the following sections.

\subsection{Rewriting Abstract Type Member}
Stainless output:
\begin{lstlisting}
[ Error  ] CurrencyUnits.scala:5:3: Stainless doesn't support abstract type members
           type A
\end{lstlisting}
This should be easy to rewrite by using generics instead of an abstract type, right?
Unfortunately not.
The problem is, CurrencyUnit uses its implementing class Satoshis.

Simplified code.
\begin{lstlisting}
  sealed abstract class CurrencyUnit {
    def +(c: CurrencyUnit): CurrencyUnit =
      Satoshis(satoshis.underlying + c.satoshis.underlying)
  }

  sealed abstract class Satoshis extends CurrencyUnit
\end{lstlisting}
Thus, when we use generics, Satoshis is not of type CurrencyUnit anymore and neither of type CurrencyUnit[A], where A is the new generic type.
Satoshis extends CurrencyUnit with type Int64, so it is of type CurrencyUnit[Int64].
That's to specific.

Since there is no easy way to fix it and the code should stay as much as possible the original, we decided to remove the abstract type.
The verification is a bit more limited, to only Satoshis, but that's no problem, because the goal is to verify the addition of Satoshis.

So by removing the abstract type and fix it to Int64 as defined by Satoshis, there was one error less in Stainless.

\subsection{Rewriting Generics}
Stainless output:
\begin{lstlisting}
  [ Error  ] NumberType.scala:3:30: Unknown type parameter type T
             sealed abstract class Number[T <: Number[T]]

\end{lstlisting}

\subsection{Rewriting Objects}
// TODO reason? because we can not extend case object?
Stainless output:
\begin{lstlisting}
[ Error  ] CurrencyUnits.scala:54:1: Objects cannot extend classes or implement
           traits, use a case object instead
           object Satoshis extends BaseNumbers[Satoshis] {
\end{lstlisting}

\subsection{Rewriting BigInt Constructor (only literal argument, no long argument, etc.)}
Stainless output:
\begin{lstlisting}
[ Error  ] CurrencyUnits.scala:47:33: Only literal arguments are allowed for BigInt.
           def toBigInt: BigInt = BigInt(toLong)
\end{lstlisting}
As described before, Stainless supports only a subset of Scala.
They provide their own dedicated BigInt library.
This library does not have support for dynamic BigInt construction only for string literals.

Again, a simplified version.
\begin{lstlisting}
  sealed abstract class Satoshis extends CurrencyUnit {
    def toBigInt: BigInt = BigInt(toLong)
    def toLong: Long = underlying.toLong
  }

  object Int64 extends BaseNumbers[Int64] {
    private case class Int64Impl(underlying: BigInt) extends Int64
  }
\end{lstlisting}
This would be really hard to refactor, because Bitcoin-S-Core tries to be as much dynamic as possible so it can be used with other cryptocurrencies too.
Maybe it would even be impossible, because they need to parse a lot from the bitcoin network.

Restricting the code one more time helps.
\emph{toBigInt} can directly return underlying, because in case of Int64, which is what Satoshis use, it is a BigInt.
So it does not be converted from BigInt to Long and back to BigInt.

\subsection{Rewriting Usage of BigInt \&-Function}

\subsection{Rewriting require (funktioniert evtl?)}