### 2019-05-25 doc/chapters/towards_verifying.tex
splitted chapter

\chapter{Trying to Verify checkTransaction}
\label{chap:connecting}

\section{Integration (Versionskonflikte, neues Plugin)}

\section{Error Reporting with sbt and JAR}

\section{Findings}

\section{Bugfix}