### 2019-05-28 doc/chapters/conclusion.tex
removed placeholder

\chapter{Conclusion}
\label{chap:conclusion}

\section{Future Work}