### 2019-06-08 doc/chapters/challenge_stainless.tex
added some text

\chapter{Practical Challenges with Stainless}
\label{chap:appendix_arb}

\section{sbt vs JAR}

We can either use the sbt plugin or a JAR file to check the code with Stainless.

Invoking the JAR on our source code Stainless will verify it.
If we have a bigger project, this becomes really tricky because we must pass all files needed including the dependencies.
This is in contrast to the sbt plugin, where we can integrate Stainless in our compilation process.
When we call compile, Stainless verifies the code and stops the compilation if the verification fails.

Having a fixed version configured in the sbt build file, every developer has the same Stainless features available.
This should prevent incompatibility with new or deprecated features when we use different plugins.

So the sbt plugin has clear advantages over the JAR file since its integrated directly.
We do not have to download it manually and find the right version and if we bump the version we can just edit it in the build file and every developer is on the same version again.

However, there are some drawbacks currently.
For example the sbt plugin does not always report errors.
We will read mor about that shortly.

\section{Integration}

During this work, Stainless updated the sbt plugin to support sbt 1.2.8 from 0.13.17 and Scala 2.12.8 from 2.11.12.
So this section might be out of date now.

There are two ways to integrate Stainless in a Scala project, the Scala build tool (sbt) plugin or a command line tool.
Both, when run, analyses the code passed and report warnings to the console about the given code.
Stainless requires and Scala recommends Java SE Development Kit 8.
Newer Java versions won't work.

To use the latest version of the sbt tool you have to build it locally.
You can run \code{sbt universal:stage} in the cloned Stainless git repository.
This generates \emph{frontends/scalac/target/universal/stage/bin/stainless-scalac}.

Bitcoin-S-Core uses sbt 1.2.8 and Scala 2.12.8, while Stainless sbt plugin is on sbt 0.13.17 and Scala 2.11.12.

Sbt introduced new features in the 1.x release used by Bitcoin-S.
Most of them can be written the sbt 0.13.17 way.

The bigger problem is, due to the different Scala and sbt versions, the error message, after trying to go in a sbt shell.
\begin{verbatim}
  [warn] There may be incompatibilities among your library dependencies; run 'evicted'
         to see detailed eviction warnings.
  [error] java.lang.NoClassDefFoundError: sbt/SourcePosition
  ...
  Project loading failed: (r)etry, (q)uit, (l)ast, or (i)gnore?
\end{verbatim}

Downgrading Bitcoin-S sbt version to 0.13.17 fixes the error but then it can not load some libraries only compiled for newer versions.
So this would take too much time to fix and changes the Bitcoin-S code unwantingly.

The next approach is to use stainless cli instead of sbt.
Running stainless on all source files does not work, because the dependencies are missing.
The parameter \emph{-classpath} can resolve it but the value of this parameter must be the paths to all the dependencies separated by a ':'.
Finally, \emph{core} depends on \emph{secp256k1jni}, another package of Bitcoin-S written in Java.
So this needs to be in the source files to.

The final command looks like this in \emph{core} folder of Bitcoin-S:
\begin{lstlisting}[language=bash]
  $ stainless
    -classpath ".:$(find ~/.ivy2/ -type f -name *.jar | tr '\n' ':')"
    $(find . -type f -name *.scala | tr '\n' ' ')
    $(find ../secp256k1jni -type f -name *.java | tr '\n' ' ')
\end{lstlisting}

The output was some Stainless error.
\begin{verbatim}
  [Internal] Error: object scala.reflect.macros.internal.macroImpl in compiler mirror
             not found.. Trace:
  [Internal] - scala.reflect.internal.MissingRequirementError$.signal
             (MissingRequirementError.scala:17)
  ...
  [Internal] object scala.reflect.macros.internal.macroImpl in compiler mirror not found.
  [Internal] Please inform the authors of Inox about this message
\end{verbatim}

After some discussion, we decided to go another way, because the errors may take too much time and it might lead to a next error.
Let's extract the code needed to verify a transaction mainly the class Transaction and ScriptInterpreter with many other classes their depending on.

After this extraction Stainless was successfully integrated with both, sbt and JAR.
This leads us to another finding as described in the next section.

\section{Error Reporting with sbt and JAR}

Running \code{sbt compile} in the project with Stainless ended without error.
But it also ended with no output.
So we were not able to change the code so Stainless would accept it.

After running the JAR file on the code there were many errors as described in chapter \ref{chap:verify_add}.

So the sbt plugin does not complain always but the JAR file did.
The open issue \mypound484 on GitHub\footnote{url{https://github.com/epfl-lara/stainless/issues/484}} might describe this error.