### 2019-06-12 doc/chapters/conclusion.tex

\chapter{Conclusion}
\label{chap:conclusion}

As we saw, it is really hard to verify software not written with verification in mind.
It needs a lot of changes in the software because verification is mathematical and the current software is written mostly in object-oriented style.
Software written in the functional paradigm would be much easier to reason about.

Thus, either Stainless must find ways to translate more of built-in object-oriented patterns of Scala to their verification tool or developers must invest more in functional programming.

The code we verified in the previous chapter is not the original code from Bitcoin-S.
So we can not guarantee the correctness of the addition of Satoshis with zero in Bitcoin-S.
Not all changes we made were as trivial as the replacement of objects with case objects.
For these non-trivial changes, as seen in section \ref{sec:bound_check}, we cannot say whether they are equivalent to the original implementation or not.